---
title: "Processing 2020_10_stations log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic station data 2020 October

#### 1. Load libraries and set current working directory
```{r cars}
library(tidyverse)
library(readr)
library(dbplyr)
library(filesstrings)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data")
input_file <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_station_data/2020_10_stations.csv"
processed_stations_file <- "2020_10_stations_processed.csv"
processed_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data"
```

#### 2. Load libraries and set current working directory
```{r}
stations_data <- read_csv(input_file)
glimpse(stations_data)
```
#### 3. Check number of station records
```{r}
id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()

glimpse(id_name)

id_name %>% count(station_id)

id_name %>% count(station_name)
```
#### Observations:
* 670 unique station_id, station_name records
* 666 unique station ids
* 670 unique station names

```{r}
ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)
filter(id_name, station_id == "26")

filter(id_name, station_id == "503")
filter(id_name, station_id == "725")
filter(id_name, is.na(station_id))
```
#### Observations:
* For stations "McClurg Ct & Illinois St" and "New St & Illinois St" given the same station_id "26".
* For stations "Drake Ave & Fullerton Ave" and "St. Louis Ave & Fullerton Ave" given the same station_id "503".
* For stations "Western Ave & 104th St" and "Halsted St & 104th St" given the same station_id "725".
* For stations "W Armitage Ave & N Sheffield Ave", and "W Oakdale Ave & N Broadway" station_id was not assigned.(same issue appear in 2020_11_stations data set)

#### 4. Correct station_name value:
```{r}
stations_data %>% filter(station_id == "26")
stations_data %>% filter(station_name == "McClurg Ct & Illinois St")
stations_data %>% filter(station_name == "New St & Illinois St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "McClurg Ct & Illinois St", "26-MCCLURG-CT-ILLINOISS"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "New St & Illinois St", "26-NEW-STR-ILLINOIS-STR"))
stations_data %>% filter(station_name == "McClurg Ct & Illinois St")
stations_data %>% filter(station_name == "New St & Illinois St")

stations_data %>% filter(station_id == "503")
stations_data %>% filter(station_name == "Drake Ave & Fullerton Ave")
stations_data %>% filter(station_name == "St. Louis Ave & Fullerton Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Drake Ave & Fullerton Ave", "503-DRAKE-AVE-FULLERTON"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "St. Louis Ave & Fullerton Ave", "503-ST-LOUIS-AVE-FULLER"))
stations_data %>% filter(station_name == "Drake Ave & Fullerton Ave")
stations_data %>% filter(station_name == "St. Louis Ave & Fullerton Ave")

stations_data %>% filter(station_id == "725")
stations_data %>% filter(station_name == "Western Ave & 104th St")
stations_data %>% filter(station_name == "Halsted St & 104th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Western Ave & 104th St", "725-WESTERN-AVE-104TH-S"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Halsted St & 104th St", "725-HALSTED-ST-104TH-ST"))
stations_data %>% filter(station_name == "Western Ave & 104th St")
stations_data %>% filter(station_name == "Halsted St & 104th St")

stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Armitage Ave & N Sheffield Ave", "TEMPORY_STATION_ID_0002"))
stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")

stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Oakdale Ave & N Broadway", "TEMPORY_STATION_ID_0003"))
stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")


id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()
glimpse(id_name)
id_name %>% count(station_id)
id_name %>% count(station_name)
```
#### Observations:
* For stations W Armitage Ave & N Sheffield Ave", and "W Oakdale Ave & N Broadway" a temporary station_id was assigned.(ids same as 2020_11_stations data set)
* 670 unique records
* unique station_id count equals to the unique_station_name count (670)

#### 5. Check for station_id consistency
```{r}
id_name <- id_name %>% mutate(id_length = nchar(station_id))
glimpse(id_name)
id_name %>% select(id_length) %>% unique()
```
#### Observations:
* There are 4 different station_id lengths

```{r}
id_name %>% filter(id_length == 3)
id_name %>% filter(id_length == 2)
id_name %>% filter(id_length == 1)
id_name %>% filter(id_length == 23)
```
#### Observations:
* Station ids with length of 23 need to be corrected with other data sets.

#### 6. Arrange dataset:
```{r}
glimpse(stations_data)
stations_data2 <- stations_data
stations_data2 <- stations_data2 %>% unique()
stations_data2 <- stations_data2 %>% arrange(latitude, longitude)
glimpse(stations_data2)
```
##### 7. Organize files
```{r}
write_csv(stations_data2, processed_stations_file)
file.move(processed_stations_file, processed_data_path , overwrite = TRUE)
```