---
title: "Processing 2020_11_stations log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic station data 2020 November

#### 1. Load libraries and set current working directory
```{r cars}
library(tidyverse)
library(readr)
library(dbplyr)
library(filesstrings)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data")
input_file <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_station_data/2020_11_stations.csv"
processed_stations_file <- "2020_11_stations_processed.csv"
processed_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data"
```

#### 2. Load libraries and set current working directory
```{r}
stations_data <- read_csv(input_file)
glimpse(stations_data)
```
#### 3. Check number of station records
```{r}
id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()

glimpse(id_name)

id_name %>% count(station_id)

id_name %>% count(station_name)
```
#### Observations:
* 666 unique station_id, station_name records
* 663 unique station ids
* 666 unique station names

```{r}
ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)
filter(id_name, station_id == "625")

stations_data %>% filter(station_id == "625")

stations_data %>% filter(station_name == "Dodge Ave & Main St")
stations_data %>% filter(station_name == "Chicago Ave & Dempster St")

stations_data %>% filter(is.na(station_id))
stations_data %>% filter(station_name == "S Michigan Ave & E 118th St")
stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")
stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")
```
#### Observations:
* Stations "Chicago Ave & Dempster St" and "Dodge Ave & Main St" given the same station_id 625. (This same issue appear in 2020_12_stations data set as well)
* Station name "Dodge Ave & Main St" appear only once with station_id 625.
* Station name "Chicago Ave & Dempster St" appear many times with station_id 625. 
* For stations "S Michigan Ave & E 118th St", "W Armitage Ave & N Sheffield Ave", and "W Oakdale Ave & N Broadway" station_id was not assinged.

#### 4. Correct station_name value:
```{r}
stations_data %>% filter(station_id == "625")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Dodge Ave & Main St", "E011"))
stations_data %>% filter(station_id == "E011")
stations_data %>% filter(station_id == "625")

stations_data %>% filter(station_name == "S Michigan Ave & E 118th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "S Michigan Ave & E 118th St", "TEMPORY_STATION_ID_0001"))
stations_data %>% filter(station_name == "S Michigan Ave & E 118th St") 

stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Armitage Ave & N Sheffield Ave", "TEMPORY_STATION_ID_0002"))
stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")

stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Oakdale Ave & N Broadway", "TEMPORY_STATION_ID_0003"))
stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")


id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()
glimpse(id_name)
id_name %>% count(station_id)
id_name %>% count(station_name)
```
#### Observations:
* station_id for "Dodge Ave & Main St" assigned as "E011" (recorded from 2021_12_stations data set)
* For stations "S Michigan Ave & E 118th St", "W Armitage Ave & N Sheffield Ave", and "W Oakdale Ave & N Broadway" a temporary station_id was assigned.
* 666 unique records
* unique station_id count equals to the unique_station_name count (666)

#### 5. Check for station_id consistency
```{r}
id_name <- id_name %>% mutate(id_length = nchar(station_id))
glimpse(id_name)
id_name %>% select(id_length) %>% unique()
```
#### Observations:
* There are 5 different station_id lengths

```{r}
id_name %>% filter(id_length == 3)
id_name %>% filter(id_length == 2)
id_name %>% filter(id_length == 1)
id_name %>% filter(id_length == 4)
id_name %>% filter(id_length == 23)
```
#### Observations:
* 3 temporary station ids need to be corrected from other data sets.

#### 6. Arrange dataset:
```{r}
glimpse(stations_data)
stations_data2 <- stations_data
stations_data2 <- stations_data2 %>% unique()
stations_data2 <- stations_data2 %>% arrange(latitude, longitude)
glimpse(stations_data2)
```

##### 7. Organize files
```{r}
write_csv(stations_data2, processed_stations_file)
file.move(processed_stations_file, processed_data_path , overwrite = TRUE)
```

