---
title: "Processing 2021_08_stations log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic station data 2021 August

#### 1. Load libraries and set current working directory
```{r cars}
library(tidyverse)
library(readr)
library(dbplyr)
library(filesstrings)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data")
input_file <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_station_data/2021_08_stations.csv"
processed_stations_file <- "2021_08_stations_processed.csv"
processed_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data"
```

#### 2. Load libraries and set current working directory
```{r}
stations_data <- read_csv(input_file)
glimpse(stations_data)
```
#### 3. Check number of station records
```{r}
id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()

glimpse(id_name)
# 729 unique records

id_name %>% count(station_id)
# 718 station ids

id_name %>% count(station_name)
```
#### Observations:
* 729 unique station_id, station_name records
* 728 unique station ids
* 729 unique station names.

```{r}
ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)
filter(id_name, station_id == "351")
```
#### Observations:
* station_id 351 appear more than once
* station_id entered as station_name in one entry.

#### 4. Correct station_name value:
```{r}
stations_data %>% filter(station_id == "351")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "351", "Mulligan Ave & Wellington Ave"))
stations_data %>% filter(station_id == "351")

id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()
glimpse(id_name)
id_name %>% count(station_id)
id_name %>% count(station_name)
```
#### Observations:
* 728 unique records
* unique station_id count equals to the unique_station_name count (728)

#### 5. Check for station_id consistency
```{r}
id_name <- id_name %>% mutate(id_length = nchar(station_id))
glimpse(id_name)
id_name %>% select(id_length) %>% unique()
```
#### Observations:
* There are 9 different station_id lengths

```{r}
id_name %>% filter(id_length == 5)
id_name %>% filter(id_length == 4)
id_name %>% filter(id_length == 6)
id_name %>% filter(id_length == 7)
id_name %>% filter(id_length == 8)
id_name %>% filter(id_length == 3)
id_name %>% filter(id_length == 35)
id_name %>% filter(id_length == 12)
id_name %>% filter(id_length == 10)
```
#### Observations:
* Following station_id have in-consistence naming
   + 20.0
   + 201022
   + 202480.0
   + 20246.0			
   + 20247.0
   + 20249.0
   + 20251.0
   + 20252.0
   + 20253.0
   + 20254.0
   + 20256.0
   + 20257.0
   + 329.0
   + 330.0
   + 331.0
   + 332.0
   + 334.0
   + 335.0
   + 362.0
   + 364.0
   + 365.0
   + 366.0
   + 368.0
   + Hubbard Bike-checking(LBS-WH-TEST)
      
#### Correct station_id values:
```{r}
remove(stations_data2)
stations_data2 <- stations_data
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20.0", "20"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "201022", "20102"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "202480.0", "20248"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20246.0", "20246"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20247.0", "20247"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20249.0", "20249"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20251.0", "20251"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20252.0", "20252"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20253.0", "20253"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20254.0", "20254"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20256.0", "20256"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20257.0", "20257"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "329.0", "329"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "330.0", "330"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "331.0", "331"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "332.0", "332"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "334.0", "334"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "335.0", "335"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "362.0", "362"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "363.0", "363"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "364.0", "364"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "365.0", "365"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "366.0", "366"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "368.0", "368"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "Hubbard Bike-checking (LBS-WH-TEST)", "LBS-WH-TEST"))
```

#### re-check station_id lengths
```{r}
remove(id_name2)
id_name2 <- stations_data2  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()
id_name2 <- id_name2 %>% mutate(id_length = nchar(station_id))
glimpse(id_name2)
id_name2 %>% select(id_length) %>% unique()
```
#### Observations:
* Number of different station_id lengths reduced from 9 to 8

```{r}
id_name2 %>% filter(id_length == 5)
id_name2 %>% filter(id_length == 2)
id_name2 %>% filter(id_length == 3)
id_name2 %>% filter(id_length == 4)
id_name2 %>% filter(id_length == 12)
id_name2 %>% filter(id_length == 10)
id_name2 %>% filter(id_length == 11)
id_name2 %>% filter(id_length == 6)
```
#### Observations:
* Station_ids are more consistent.

#### 6. Arrange dataset:
```{r}
glimpse(stations_data2)
stations_data2 <- stations_data2 %>% unique()
stations_data2 <- stations_data2 %>% arrange(latitude, longitude)
glimpse(stations_data2)
```

##### 7. Organize files
```{r}
write_csv(stations_data2, processed_stations_file)
file.move(processed_stations_file, processed_data_path , overwrite = TRUE)
```