---
title: "Processing 2021_06_stations log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic station data 2021 June

#### 1. Load libraries and set current working directory
```{r cars}
library(tidyverse)
library(readr)
library(dbplyr)
library(filesstrings)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data")
input_file <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_station_data/2021_06_stations.csv"
processed_stations_file <- "2021_06_stations_processed.csv"
processed_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data"
```

#### 2. Load libraries and set current working directory
```{r}
stations_data <- read_csv(input_file)
glimpse(stations_data)
```
#### 3. Check number of station records
```{r}
id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()

glimpse(id_name)

id_name %>% count(station_id)

id_name %>% count(station_name)
```
#### Observations:
* 690 unique station_id, station_name records
* 690 unique station ids
* 690 unique station names

```{r}
ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)
```
#### 4. Check for station_id consistency
```{r}
id_name <- id_name %>% mutate(id_length = nchar(station_id))
glimpse(id_name)
id_name %>% select(id_length) %>% unique()
```
#### Observations:
* There are 10 different station_id lengths

```{r}
id_name %>% filter(id_length == 5)
id_name %>% filter(id_length == 4)
id_name %>% filter(id_length == 6)
id_name %>% filter(id_length == 7)
id_name %>% filter(id_length == 8)
id_name %>% filter(id_length == 3)
id_name %>% filter(id_length == 9)
id_name %>% filter(id_length == 35)
id_name %>% filter(id_length == 12)
id_name %>% filter(id_length == 10)
```
#### Observations:
* Following station_id have in-consistence naming
+ 329.0
+ 330.0
+ 331.0
+ 362.0
+ 363.0
+ 364.0
+ 365.0
+ 366.0
+ 368.0
+ 20.0
+ 20246.0	
+ 20247.0		
+ 20249.0	
+ 20251.0	
+ 20252.0		
+ 20253.0	
+ 20254.0		
+ 20256.0	
+ 20257.0
+ 202480.0
+ DIVVY CASSETTE REPAIR MOBILE STATION
+ Hubbard Bike-checking (LBS-WH-TEST)

#### Correct station_id values:
```{r}
remove(stations_data2)
stations_data2 <- stations_data
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20.0", "20"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20246.0", "20246"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20247.0", "20247"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20249.0", "20249"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20251.0", "20251"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20252.0", "20252"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20253.0", "20253"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20254.0", "20254"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20256.0", "20256"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "20257.0", "20257"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "202480.0", "202480"))
stations_data2 <- stations_data2 %>% mutate(station_id = replace(station_id, station_id == "Hubbard Bike-checking (LBS-WH-TEST)", "LBS-WH-TEST"))
```

#### re-check station_id lengths
```{r}
remove(id_name2)
id_name2 <- stations_data2  %>%  select(station_id, station_name) %>% arrange(station_id) %>% unique()
id_name2 <- id_name2 %>% mutate(id_length = nchar(station_id))
glimpse(id_name2)
id_name2 %>% select(id_length) %>% unique()
```
#### Observations:
* Number of different station_id lengths reduced from 10 to 9

```{r}
id_name2 %>% filter(id_length == 5)
id_name2 %>% filter(id_length == 2)
id_name2 %>% filter(id_length == 6)
id_name2 %>% filter(id_length == 3)
id_name2 %>% filter(id_length == 9)
id_name2 %>% filter(id_length == 4)
id_name2 %>% filter(id_length == 12)
id_name2 %>% filter(id_length == 10)
id_name2 %>% filter(id_length == 10)
```
#### Observations:
* Station_ids are more consistent.

#### 5. Arrange dataset:
```{r}
glimpse(stations_data2)
stations_data2 <- stations_data2 %>% unique()
stations_data2 <- stations_data2 %>% arrange(latitude, longitude)
glimpse(stations_data2)
```
##### 7. Organize files
```{r}
write_csv(stations_data2, processed_stations_file)
file.move(processed_stations_file, processed_data_path , overwrite = TRUE)
```