---
title: "Merge stations data log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic station data

#### 1. Load libraries and set current working directory
```{r cars}
library(tidyverse)
library(readr)
library(dbplyr)
library(filesstrings)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data")
processed_stations_file <- "stations_processed_final.csv"
processed_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/processing_data/processed_data/processed_stations_data"
```

#### 2. Merge 2021 August to 2020 September station data
```{r}
# 2021 August
input_file <- read_csv("2021_08_stations_processed.csv")
glimpse(input_file)
stations_data <- input_file

# 2021 July
input_file <- read_csv("2021_07_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 June
input_file <- read_csv("2021_06_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 May
input_file <- read_csv("2021_05_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 April
input_file <- read_csv("2021_04_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 March
input_file <- read_csv("2021_03_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 February
input_file <- read_csv("2021_02_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2021 January
input_file <- read_csv("2021_01_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2020 December
input_file <- read_csv("2020_12_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2020 November
input_file <- read_csv("2020_11_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2020 October
input_file <- read_csv("2020_10_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

# 2020 September
input_file <- read_csv("2020_09_stations_processed.csv")
glimpse(input_file)
stations_data <- rbind(stations_data, input_file)

glimpse(stations_data)
stations_data <- stations_data %>% distinct()
glimpse(stations_data)
```
#### 3. Check number of station records
```{r}
id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% distinct()

glimpse(id_name)

id_name %>% count(station_id)

id_name %>% count(station_name)
```
#### Observations:
* 1,351 unique station_id, station_name records
* 1,301 unique station ids
* 754 unique station names

```{r}
ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)
```
#### Observations:
* 47 station ids appear more than once

#### 4. Correct station_name value:
```{r}
stations_data %>% filter(station_id == "13074")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Broadway & Wilson - Truman College Vaccination Site", "Broadway & Wilson Ave"))
stations_data %>% filter(station_id == "13074")

stations_data %>% filter(station_id == "13099")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Halsted St & 18th St (Temp)", "Halsted St & 18th St"))
stations_data %>% filter(station_id == "13099")

stations_data %>% filter(station_id == "13300")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & Monroe St", "Lake Shore Dr & Monroe St"))
stations_data %>% filter(station_id == "13300")

stations_data %>% filter(station_id == "20")
stations_data %>% filter(station_name == "Damen Ave & Wabansia Ave")
stations_data %>% filter(station_name == "Sheffield Ave & Kingsbury St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Sheffield Ave & Kingsbury St", "13154"))
stations_data %>% filter(station_name == "Damen Ave & Wabansia Ave")
stations_data %>% filter(station_name == "Sheffield Ave & Kingsbury St")

stations_data %>% filter(station_id == "20248")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "N Hampden Ct & W Diversey Ave", "Hampden Ct & Diversey Ave"))
stations_data %>% filter(station_id == "20248")

stations_data %>% filter(station_id == "26")
stations_data %>% filter(station_name == "New St & Illinois St")
stations_data %>% filter(station_name == "McClurg Ct & Illinois St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "New St & Illinois St", "TA1306000013"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "McClurg Ct & Illinois St", "26"))
stations_data %>% filter(station_name == "New St & Illinois St")
stations_data %>% filter(station_name == "McClurg Ct & Illinois St")

stations_data %>% filter(station_id == "310")
stations_data %>% filter(station_name == "North Ave & New England Ave")
stations_data %>% filter(station_name == "Damen Ave & Charleston St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Damen Ave & Charleston St", "13288"))
stations_data %>% filter(station_name == "North Ave & New England Ave")
stations_data %>% filter(station_name == "Damen Ave & Charleston St")

stations_data %>% filter(station_id == "312")
stations_data %>% filter(station_name == "Lockwood Ave & Wrightwood Ave")
stations_data %>% filter(station_name == "Clarendon Ave & Gordon Ter")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Clarendon Ave & Gordon Ter", "13379"))
stations_data %>% filter(station_name == "Lockwood Ave & Wrightwood Ave")
stations_data %>% filter(station_name == "Clarendon Ave & Gordon Ter")

stations_data %>% filter(station_id == "313")
stations_data %>% filter(station_name == "Long Ave & Belden Ave")
stations_data %>% filter(station_name == "Lakeview Ave & Fullerton Pkwy")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Lakeview Ave & Fullerton Pkwy", "TA1309000019"))
stations_data %>% filter(station_name == "Long Ave & Belden Ave")
stations_data %>% filter(station_name == "Lakeview Ave & Fullerton Pkwy")

stations_data %>% filter(station_id == "316")
stations_data %>% filter(station_name == "Lamon Ave & Belmont Ave")
stations_data %>% filter(station_name == "Damen Ave & Sunnyside Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Damen Ave & Sunnyside Ave", "TA1309000012"))
stations_data %>% filter(station_name == "Lamon Ave & Belmont Ave")
stations_data %>% filter(station_name == "Damen Ave & Sunnyside Ave")

stations_data %>% filter(station_id == "317")
stations_data %>% filter(station_name == "Long Ave & Belmont Ave")
stations_data %>% filter(station_name == "Wood St & Taylor St (Temp)")
stations_data %>% filter(station_name == "Wood St & Taylor St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Wood St & Taylor St (Temp)", "13285"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Wood St & Taylor St", "13285"))
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Wood St & Taylor St (Temp)", "Wood St & Taylor St"))
stations_data %>% filter(station_name == "Long Ave & Belmont Ave")
stations_data %>% filter(station_name == "Wood St & Taylor St (Temp)")
stations_data %>% filter(station_name == "Wood St & Taylor St")

stations_data %>% filter(station_id == "319")
stations_data %>% filter(station_name == "Roscoe & Harlem")
stations_data %>% filter(station_name == "Greenview Ave & Diversey Pkwy")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Greenview Ave & Diversey Pkwy", "13294"))
stations_data %>% filter(station_name == "Roscoe & Harlem")
stations_data %>% filter(station_name == "Greenview Ave & Diversey Pkwy")

stations_data %>% filter(station_id == "321")
stations_data %>% filter(station_name == "Kostner Ave & Wrightwood Ave")
stations_data %>% filter(station_name == "Wabash Ave & 9th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Wabash Ave & 9th St", "TA1309000010"))
stations_data %>% filter(station_name == "Kostner Ave & Wrightwood Ave")
stations_data %>% filter(station_name == "Wabash Ave & 9th St")

stations_data %>% filter(station_id == "322")
stations_data %>% filter(station_name == "Cicero Ave & Grace St")
stations_data %>% filter(station_name == "Kimbark Ave & 53rd St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Kimbark Ave & 53rd St", "TA1309000037"))
stations_data %>% filter(station_name == "Cicero Ave & Grace St")
stations_data %>% filter(station_name == "Kimbark Ave & 53rd St")

stations_data %>% filter(station_id == "324")
stations_data %>% filter(station_name == "Meade Ave & Addison St")
stations_data %>% filter(station_name == "Stockton Dr & Wrightwood Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Stockton Dr & Wrightwood Ave", "13276"))
stations_data %>% filter(station_name == "Meade Ave & Addison St")
stations_data %>% filter(station_name == "Stockton Dr & Wrightwood Ave")

stations_data %>% filter(station_id == "325")
stations_data %>% filter(station_name == "Harlem Ave & Grace St")
stations_data %>% filter(station_name == "Clark St & Winnemac Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Clark St & Winnemac Ave", "TA1309000035"))
stations_data %>% filter(station_name == "Harlem Ave & Grace St")
stations_data %>% filter(station_name == "Clark St & Winnemac Ave")

stations_data %>% filter(station_id == "327")
stations_data %>% filter(station_name == "Plainfield & Irving Park")
stations_data %>% filter(station_name == "Sheffield Ave & Webster Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Sheffield Ave & Webster Ave", "TA1309000033"))
stations_data %>% filter(station_name == "Plainfield & Irving Park")
stations_data %>% filter(station_name == "Sheffield Ave & Webster Ave")

stations_data %>% filter(station_id == "329")
stations_data %>% filter(station_name == "Central Park Ave & Douglas Blvd")
stations_data %>% filter(station_name == "Lake Shore Dr & Diversey Pkwy")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Lake Shore Dr & Diversey Pkwy", "TA1309000039"))
stations_data %>% filter(station_name == "Central Park Ave & Douglas Blvd")
stations_data %>% filter(station_name == "Lake Shore Dr & Diversey Pkwy")

stations_data %>% filter(station_id == "330")
stations_data %>% filter(station_name == "Keeler Ave & Roosevelt Rd")
stations_data %>% filter(station_name == "Lincoln Ave & Addison St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Lincoln Ave & Addison St", "TA1309000050"))
stations_data %>% filter(station_name == "Keeler Ave & Roosevelt Rd")
stations_data %>% filter(station_name == "Lincoln Ave & Addison St")

stations_data %>% filter(station_id == "331")
stations_data %>% filter(station_name == "Pulaski Rd & 21st St")
stations_data %>% filter(station_name == "Halsted St & Clybourn Ave")
# No alternative id for "Halsted St & Clybourn Ave"
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Pulaski Rd & 21st St", "Halsted St & Clybourn Ave"))
stations_data %>% filter(station_name == "Pulaski Rd & 21st St")
stations_data %>% filter(station_name == "Halsted St & Clybourn Ave")

stations_data %>% filter(station_id == "332")
stations_data %>% filter(station_name == "Harding Ave & 26th St")
stations_data %>% filter(station_name == "Burling St & Diversey Pkwy")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Burling St & Diversey Pkwy", "TA1309000036"))
stations_data %>% filter(station_name == "Harding Ave & 26th St")
stations_data %>% filter(station_name == "Burling St & Diversey Pkwy")

stations_data %>% filter(station_id == "334")
stations_data %>% filter(station_name == "Lawndale Ave & 30th St")
stations_data %>% filter(station_name == "Lake Shore Dr & Belmont Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Lake Shore Dr & Belmont Ave", "TA1309000049"))
stations_data %>% filter(station_name == "Lawndale Ave & 30th St")
stations_data %>% filter(station_name == "Lake Shore Dr & Belmont Ave")

stations_data %>% filter(station_id == "335")
stations_data %>% filter(station_name == "Komensky Ave & 31st St")
stations_data %>% filter(station_name == "Calumet Ave & 35th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Calumet Ave & 35th St", "13345"))
stations_data %>% filter(station_name == "Komensky Ave & 31st St")
stations_data %>% filter(station_name == "Calumet Ave & 35th St")

stations_data %>% filter(station_id == "351")
stations_data %>% filter(station_name == "Mulligan Ave & Wellington Ave")
stations_data %>% filter(station_name == "Cottage Grove Ave & 51st St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Cottage Grove Ave & 51st St", "TA1309000067"))
stations_data %>% filter(station_name == "Mulligan Ave & Wellington Ave")
stations_data %>% filter(station_name == "Cottage Grove Ave & 51st St")

stations_data %>% filter(station_id == "352")
stations_data %>% filter(station_name == "Sayre & Diversey")
stations_data %>% filter(station_name == "Jeffery Blvd & 67th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Jeffery Blvd & 67th St", "KA1503000030"))
stations_data %>% filter(station_name == "Sayre & Diversey")
stations_data %>% filter(station_name == "Jeffery Blvd & 67th St")

stations_data %>% filter(station_id == "353")
stations_data %>% filter(station_name == "Meade Ave & Diversey Ave")
stations_data %>% filter(station_name == "Clark St & Touhy Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Clark St & Touhy Ave", "RP-007"))
stations_data %>% filter(station_name == "Meade Ave & Diversey Ave")
stations_data %>% filter(station_name == "Clark St & Touhy Ave")

stations_data %>% filter(station_id == "354")
stations_data %>% filter(station_name == "Parkside Ave & Armitage Ave")
stations_data %>% filter(station_name == "Sheridan Rd & Greenleaf Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Sheridan Rd & Greenleaf Ave", "KA1504000159"))
stations_data %>% filter(station_name == "Parkside Ave & Armitage Ave")
stations_data %>% filter(station_name == "Sheridan Rd & Greenleaf Ave")

stations_data %>% filter(station_id == "355")
stations_data %>% filter(station_name == "Central Ave & Parker Ave")
stations_data %>% filter(station_name == "South Shore Dr & 67th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "South Shore Dr & 67th St", "KA1503000029"))
stations_data %>% filter(station_name == "Central Ave & Parker Ave")
stations_data %>% filter(station_name == "South Shore Dr & 67th St")

stations_data %>% filter(station_id == "356")
stations_data %>% filter(station_name == "Lavergne & Fullerton")
stations_data %>% filter(station_name == "Stony Island Ave & 71st St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Stony Island Ave & 71st St", "KA1503000045"))
stations_data %>% filter(station_name == "Lavergne & Fullerton")
stations_data %>% filter(station_name == "Stony Island Ave & 71st St")

stations_data %>% filter(station_id == "359")
stations_data %>% filter(station_name == "Kilbourn & Belden")
stations_data %>% filter(station_name == "Larrabee St & Division St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Larrabee St & Division St", "KA1504000079"))
stations_data %>% filter(station_name == "Kilbourn & Belden")
stations_data %>% filter(station_name == "Larrabee St & Division St")

stations_data %>% filter(station_id == "364")
stations_data %>% filter(station_name == "Homan Ave & Fillmore St")
stations_data %>% filter(station_name == "Larrabee St & Oak St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Larrabee St & Oak St", "KA1504000116"))
stations_data %>% filter(station_name == "Homan Ave & Fillmore St")
stations_data %>% filter(station_name == "Larrabee St & Oak St")

stations_data %>% filter(station_id == "365")
stations_data %>% filter(station_name == "Kildare Ave & 26th St")
stations_data %>% filter(station_name == "Halsted St & North Branch St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Halsted St & North Branch St", "KA1504000117"))
stations_data %>% filter(station_name == "Kildare Ave & 26th St")
stations_data %>% filter(station_name == "Halsted St & North Branch St")

stations_data %>% filter(station_id == "366")
stations_data %>% filter(station_name == "Spaulding Ave & 16th St")
stations_data %>% filter(station_name == "Loomis St & Archer Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Loomis St & Archer Ave", "KA1503000040"))
stations_data %>% filter(station_name == "Spaulding Ave & 16th St")
stations_data %>% filter(station_name == "Loomis St & Archer Ave")

stations_data %>% filter(station_id == "368")
stations_data %>% filter(station_name == "Tripp Ave & 31st St")
stations_data %>% filter(station_name == "Ashland Ave & Archer Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Ashland Ave & Archer Ave", "15442"))
stations_data %>% filter(station_name == "Tripp Ave & 31st St")
stations_data %>% filter(station_name == "Ashland Ave & Archer Ave")

stations_data %>% filter(station_id == "393")
stations_data %>% filter(station_name == "Oketo Ave & Addison St")
stations_data %>% filter(station_name == "Calumet Ave & 71st St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Calumet Ave & 71st St", "15599"))
stations_data %>% filter(station_name == "Oketo Ave & Addison St")
stations_data %>% filter(station_name == "Calumet Ave & 71st St")

stations_data %>% filter(station_id == "396")
stations_data %>% filter(station_name == "Central Ave & Roscoe St")
stations_data %>% filter(station_name == "Yates Blvd & 75th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Yates Blvd & 75th St", "KA1503000024"))
stations_data %>% filter(station_name == "Central Ave & Roscoe St")
stations_data %>% filter(station_name == "Yates Blvd & 75th St")

stations_data %>% filter(station_id == "398")
stations_data %>% filter(station_name == "Long & Irving Park")
stations_data %>% filter(station_name == "Rainbow Beach")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Rainbow Beach", "KA1503000049"))
stations_data %>% filter(station_name == "Long & Irving Park")
stations_data %>% filter(station_name == "Rainbow Beach")

stations_data %>% filter(station_id == "631")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Malcolm X College Vaccination Site", "Malcolm X College"))
stations_data %>% filter(station_id == "631")

stations_data %>% filter(station_id == "KA1503000055")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Halsted & 63rd - Kennedy-King Vaccination Site", "Halsted St & 63rd St"))
stations_data %>% filter(station_id == "KA1503000055")

stations_data %>% filter(station_id == "KA1504000168")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Western & 28th - Velasquez Institute Vaccination Site", "Western Ave & 28th St"))
stations_data %>% filter(station_id == "KA1503000055")

stations_data %>% filter(station_id == "LF-005")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & North Blvd", "Lake Shore Dr & North Blvd"))
stations_data %>% filter(station_id == "LF-005")

stations_data %>% filter(station_id == "TA1305000039")
stations_data %>% filter(station_name == "Elston Ave & Cortland St")
stations_data %>% filter(station_name == "Marshfield Ave & Cortland St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Elston Ave & Cortland St", "TA1305000039"))
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Marshfield Ave & Cortland St", "58"))
stations_data %>% filter(station_name == "Elston Ave & Cortland St")
stations_data %>% filter(station_name == "Marshfield Ave & Cortland St")

stations_data %>% filter(station_id == "TA1306000029")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & Ohio St", "Lake Shore Dr & Ohio St"))
stations_data %>% filter(station_id == "TA1306000029")

stations_data %>% filter(station_id == "TA1307000041")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & Wellington Ave", "Lake Shore Dr & Wellington Ave"))
stations_data %>% filter(station_id == "TA1307000041")

stations_data %>% filter(station_id == "TA1309000039")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & Diversey Pkwy", "Lake Shore Dr & Diversey Pkwy"))
stations_data %>% filter(station_id == "TA1309000039")

stations_data %>% filter(station_id == "TA1309000049")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "DuSable Lake Shore Dr & Belmont Ave", "Lake Shore Dr & Belmont Ave"))
stations_data %>% filter(station_id == "TA1309000049")

stations_data %>% filter(is.na(station_id))
stations_data %>% filter(station_name == "S Michigan Ave & E 118th St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "S Michigan Ave & E 118th St", "20209"))
stations_data %>% filter(station_name == "S Michigan Ave & E 118th St")
stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Armitage Ave & N Sheffield Ave", "20254"))
stations_data %>% filter(station_name == "W Armitage Ave & N Sheffield Ave")
stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "W Oakdale Ave & N Broadway", "20252"))
stations_data %>% filter(station_name == "W Oakdale Ave & N Broadway")
stations_data %>% filter(station_name == "Dodge Ave & Main St")
stations_data <- stations_data %>% mutate(station_id = replace(station_id, station_name == "Dodge Ave & Main St", "E011"))
stations_data %>% filter(station_name == "Dodge Ave & Main St")
```

```{r}
glimpse(stations_data)
stations_data <- stations_data %>% unique()
glimpse(stations_data)
# Records reduced from 2,209,085 to 2,208,854 

id_name <- stations_data  %>%  select(station_id, station_name) %>% arrange(station_id) %>% distinct()
glimpse(id_name)
id_name %>% count(station_id)

ids <- id_name %>% select(station_id) %>% count(station_id)
glimpse(ids)

filter(ids, n > 1)

names <- id_name %>% select(station_name) %>% count(station_name)
glimpse(names)
names %>% filter(n > 1)
```
#### Observations:
* 1,293 unique records
* 1,293 unique station ids
* 740 unique station names

```{r}
id_name2 <- id_name %>%  arrange(station_name) %>% group_by(station_name) %>% slice(1) %>% ungroup()

#View(id_name2)

id_name2 <- id_name2 %>% mutate(id_length = nchar(station_id))
id_name2 %>% select(id_length) %>% distinct()
```

```{r}
id_name2 %>% filter(id_length == 3) 
id_name2 %>% filter(id_length == 5)
id_name2 %>% filter(id_length == 11)
id_name2 %>% filter(id_length == 12)
id_name2 %>% filter(id_length == 2)
id_name2 %>% filter(id_length == 6)
id_name2 %>% filter(id_length == 36)
id_name2 %>% filter(id_length == 4)
id_name2 %>% filter(id_length == 23)
id_name2 %>% filter(id_length == 1)
id_name2 %>% filter(id_length == 9)
```

```{r}
id_name2 %>% filter(station_name == "Franklin St & Adams St (Temp)")
id_name2 <- id_name2 %>% mutate(station_name = replace(station_name, station_name == "Franklin St & Adams St (Temp)", "Franklin St & Adams St"))
id_name2 %>% filter(station_id == "286")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Franklin St & Adams St (Temp)", "Franklin St & Adams St"))

id_name2 %>% filter(station_name == "Pulaski Rd & Eddy St (Temp)")
id_name2 <- id_name2 %>% mutate(station_name = replace(station_name, station_name == "Pulaski Rd & Eddy St (Temp)", "Pulaski Rd & Eddy St"))
id_name2 %>% filter(station_id == "488")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Pulaski Rd & Eddy St (Temp)", "Pulaski Rd & Eddy St"))

id_name2 %>% filter(station_name == "Wentworth Ave & 24th St (Temp)")
id_name2 <- id_name2 %>% mutate(station_name = replace(station_name, station_name == "Wentworth Ave & 24th St (Temp)", "Wentworth Ave & 24th St"))
id_name2 %>% filter(station_id == "132")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "Wentworth Ave & 24th St (Temp)", "Wentworth Ave & 24th St"))

id_name2 %>% filter(station_name == "California Ave & Francis Pl (Temp)")
id_name2 <- id_name2 %>% mutate(station_name = replace(station_name, station_name == "California Ave & Francis Pl (Temp)", "California Ave & Francis Pl"))
id_name2 %>% filter(station_id == "13259")
stations_data <- stations_data %>% mutate(station_name = replace(station_name, station_name == "California Ave & Francis Pl (Temp)", "California Ave & Francis Pl"))

stations_data %>% filter(station_name == "Drake Ave & Fullerton Ave")
stations_data %>% filter(station_id == "503")
stations_data %>% filter(station_name == "Jeffery Blvd & 91st St")
stations_data %>% filter(station_id == "704")

id_name2 <- id_name2 %>% select(-id_length)
#View(id_name2)
# 740 stations
```
#### Assign corrected station ids:
```{r}
#joining station data 
stations_data2 <- left_join(stations_data, id_name2, by = "station_name")
View(stations_data2)

station_data_final <- stations_data2 %>% select(station_id.y, station_name, latitude, longitude)
glimpse(station_data_final)
station_data_final <- rename(station_data_final, station_id = station_id.y)

glimpse(station_data_final)

station_data_final <- station_data_final %>% distinct()
glimpse(station_data_final)
# Records 2,206,745

id_name3 <- station_data_final  %>%  select(station_id, station_name)  %>% distinct() %>% arrange(station_id)

id_name3 %>% select(station_id) %>% distinct() %>% count()
# 740
id_name3 %>% select(station_name) %>% distinct() %>% count()
# 740
```
#### Observations:
* 740 unique station ids
* 740 unique station names


#### Save final results:
```{r}
glimpse(id_name3)
View(id_name3)
View(station_data_final)

temp_coordinates <- station_data_final
View(temp_coordinates)
#temp_coordinates <- temp_coordinates %>% mutate(latitude_r3 = round(latitude, 3))
#temp_coordinates <- temp_coordinates %>% mutate(longitude_r3 = round(longitude, 3))
#View(temp_coordinates)
#temp_coordinates <- temp_coordinates %>% select(-latitude, -longitude)
temp_coordinates <- temp_coordinates %>% distinct()
temp_coordinates <- temp_coordinates %>% group_by(station_name) %>% slice(1) %>% ungroup()

#temp_coordinates <- rename(temp_coordinates, location_latitude = latitude_r3, location_longitude = longitude_r3)

View(temp_coordinates)

write_csv(temp_coordinates, "station_by_location.csv")
write_csv(station_data_final, "stations_processed_final.csv")

```