---
title: "Cleaning 202106-divvy-tripdata log"
author: "Yasith Udawatte"
date: "2021-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cyclistic trip data 2021 June

#### Load libraries and set current working directory
```{r}
library(tidyverse)
library(readr)
library(filesstrings)
library(lubridate)

setwd("D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data")
input_file <- "202106-divvy-tripdata.csv"
cleaned_tripdata_file <- "2021_06_tripdata.csv"
cleaned_station_file <- "2021_06_stations.csv"

cleaned_trip_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_trip_data"
cleaned_stations_data_path <- "D:/Documents/R/Projects/Cyclistic/Cyclistic_Bike_Share/cleaning_data/cleaned_data/cleaned_station_data"

valid_start_at <- ymd_hms("2021-06-01 00:00:00")
valid_end_at <- ymd_hms("2021-06-30 23:59:59")
```

#### Read csv 
```{r}
monthly_trip_data  <- read_csv(input_file, 
    col_types = cols(ride_id = col_character(), 
        rideable_type = col_character(), 
        started_at = col_datetime(), 
        ended_at = col_datetime(), 
        start_station_name = col_character(), 
        start_station_id = col_character(), 
        end_station_name = col_character(), 
        end_station_id = col_character(), 
        start_lat = col_double(), start_lng = col_double(), 
        end_lat = col_double(), end_lng = col_double(), 
        member_casual = col_character()))

str(monthly_trip_data)
glimpse(monthly_trip_data)
```

#### Rename columns
```{r}
monthly_trip_data <- rename(
  monthly_trip_data, 
  start_latitude = start_lat, 
  end_latitude = end_lat, 
  start_longitude = start_lng, 
  end_longitude = end_lng,
  user_type = member_casual)

glimpse(monthly_trip_data)
```

#### Check columns
##### 1. ride_id
```{r}
monthly_trip_data %>% filter(is.na(ride_id))
monthly_trip_data %>% filter(duplicated(ride_id))
```
##### Observations:
* No null values.
* No duplicate values.

##### 2. rideable_type 
```{r}
monthly_trip_data %>% filter(is.na(rideable_type))
monthly_trip_data %>% select(rideable_type) %>% unique()
```
##### Observations:
* No null values, only 3 rideable types available.
* Values are consistent.

##### 3. strated_at
```{r}
monthly_trip_data %>% filter(is.na(started_at))
monthly_trip_data %>% filter(started_at < valid_start_at | started_at > valid_end_at )
```
##### Observations:
* No null values.
* Column type is date-time.

##### 4. ended_at
```{r}
monthly_trip_data %>% filter(is.na(ended_at))
```
##### Observations:
* No null values.
* Column type is date-time.

##### 5. Add new column: duration
```{r}
monthly_trip_data <- mutate(
  monthly_trip_data, 
  duration_in_mins = difftime(ended_at, started_at, units = "mins") 
  )

glimpse(monthly_trip_data)

monthly_trip_data %>%  filter(duration_in_mins <= 0) %>% select(ride_id, started_at, ended_at, duration_in_mins) 
```
##### Observations:
* There are 66 records where started_at value is greater than or equal to the ended_at value.
* started_at >= ended_at records need to be removed from the data set.

```{r}
monthly_trip_data_v1 <- monthly_trip_data %>%  filter(duration_in_mins > 0)
glimpse(monthly_trip_data_v1)
```
##### Observations:
* monthly_trip_date_v1 contains only duration_in_seconds positive records.

##### 6. Check user_type
```{r}
monthly_trip_data_v1 %>% select(user_type) %>% unique()
```
##### Observations:
* No empty values.
* Two user types available (member, casual).

##### 7. Check stations data
```{r}
start_station_data <- monthly_trip_data_v1 %>% 
  select(start_station_name, start_station_id, start_latitude, start_longitude) %>% 
  unique()

glimpse(start_station_data)

start_station_data %>% filter(is.na(start_latitude) | is.na(start_longitude))
start_station_data %>% select(start_station_name) %>% unique() %>% count()
start_station_data %>% select(start_station_id) %>% unique() %>% count()

start_station_data <- mutate(start_station_data, start_station_id_length = nchar(start_station_id)) 
start_station_data %>% select(start_station_id_length) %>% unique() 
```
##### Start station data - Observations:
* Null values exists in start_station_name and start_station_id columns.
* Each record have both latitude and longitude values.
* The count of unique start station names is 690. 
* The count of unique start station ids is 690. 
* start_station_id naming is not consistent, different lengths and different formats.

```{r}
end_station_data <- monthly_trip_data_v1 %>% 
  select(end_station_name, end_station_id, end_latitude, end_longitude) %>% 
  unique()

glimpse(end_station_data)

end_station_data %>% filter(is.na(end_latitude) | is.na(end_longitude))

end_station_data %>% select(end_station_name) %>% unique() %>% count()
end_station_data %>% select(end_station_id) %>% unique() %>% count()

end_station_data <- mutate(end_station_data, end_station_id_length = nchar(end_station_id)) 
end_station_data %>% select(end_station_id_length) %>% unique() 

```
##### Start station data - Observations:
* Null values exists in end_station_name and end_station_id columns.
* A record exists with null values for both end_latitude and end_longitude.
* The count of unique start station names is 691.
* The count of unique start station ids is 691.
* end_station_id naming is not consistent, different lengths and different formats.

##### 8. Arrange station data
```{r}
glimpse(start_station_data)
glimpse(end_station_data)

start_station_data <- start_station_data %>% select(-start_station_id_length)
end_station_data <- end_station_data %>% select(-end_station_id_length)

start_station_data <- rename(
  start_station_data,
  station_name = start_station_name,
  station_id = start_station_id,
  latitude = start_latitude,
  longitude = start_longitude
)

end_station_data <- rename(
  end_station_data,
  station_name = end_station_name,
  station_id = end_station_id,
  latitude = end_latitude,
  longitude = end_longitude
)

# remove null station entries
start_station_data <-  start_station_data %>% filter(!is.na(station_name) | !is.na(station_id))
glimpse(start_station_data)

end_station_data <-  end_station_data %>% filter(!is.na(station_name) | !is.na(station_id))
glimpse(end_station_data)

stations <- rbind(start_station_data, end_station_data)
glimpse(stations)

stations <- stations %>% unique()
glimpse(stations)

stations <- stations %>% arrange(latitude, longitude)
glimpse(stations)
```

##### 9. Removing empty latitude and longitude records
```{r}
glimpse(monthly_trip_data_v1)
temp_rec <- monthly_trip_data_v1 %>% filter(is.na(start_latitude) | is.na(start_longitude))
glimpse(temp_rec)

monthly_trip_data_v1 %>% filter(is.na(end_latitude) | is.na(end_longitude))
monthly_trip_data_v1 %>% filter(is.na(end_latitude) | is.na(end_longitude)) %>% count()

```
##### Observations:
* There are 717 entries with null values for end_latitude and end_longitude fields


```{r}
monthly_trip_data_v2 <- monthly_trip_data_v1 %>% filter(!is.na(end_latitude) & !is.na(end_longitude))
glimpse(monthly_trip_data_v2)
```

##### 10. Organize files
```{r}
write_csv(stations, cleaned_station_file)
write_csv(monthly_trip_data_v2, cleaned_tripdata_file )
file.move(cleaned_station_file, cleaned_stations_data_path, overwrite = TRUE)
file.move(cleaned_tripdata_file, cleaned_trip_data_path, overwrite = TRUE)
```
